Report on Late Winter2013 Production: Image Differencing

### 2013-03-20 W13report.tex
Many figures.

\subsection{Algorithm}

We establish the math behind the process below.

@@

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We use variations on the
configuration where we use one Gaussian, with the above width, and
three Gaussians, the central of which has this width, and the others a
factor of $\beta = 2.0$ smaller and larger.  We modify each of the
Gaussians by a set of Laguerre polynomials of a given order.  The
smallest Gaussian is modified to the highest order, with the others
modified by order//2.  Figure~\ref{basis} provides an example basis
set with this configuration.  The total number of shapes in the basis
are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription to basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 2.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.
}.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

A template image was generated by {\bf DETAIL}.

@@

The theoretical expectations for the number of false positives due to
random Gaussian fluctuations is outlined in \cite{Kaiser-PointSources}.

@@

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 5                      & 29  \\
         & False     & 10                     & 81  \\
v6866602 & True      & 5                      & 30  \\
         & False     & 9                      & 35  \\
v6866603 & True      & 2                      & 9  \\
         & False     & 1                      & 23  \\
\end{tabular}
\caption{Total number of {\tt DiaSources} from all 9 sensors in raft
  2,2, for the configurations that yielded the lowest number of false
  positives.  Matches are determined using a 2'' match radius with the
  input reference catalog.  Objects with the {\tt EDGE} flag are
  ignored. \label{tab-bestfp}}
\end{table}

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Configs} \\
\hline
Visit    & Prefilter & nGauss & degGauss & spatialOrder \\
\hline
v6866601 & True      & 3      & 6        & 5 \\
         & False     & 3      & 2        & 4 \\
v6866602 & True      & 3      & 6        & 5 \\
         & False     & 3      & 5        & 5 \\
v6866603 & True      & 3      & 6        & 4 \\
         & False     & 3      & 6        & 4 \\
\end{tabular}
\caption{Configurations that led to the best results listed in
  Table~\ref{tab-bestfp}.  All post-production runs include
  perturbations about these configurations. \label{tab-bestconfig}}
\end{table}

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/falsePos.eps} \\
\caption{``Heat--map'' detailing the total numbers of false positives
  as a function of two main degrees of freedom: the maximum order of
  basis Laguerre polynomials along the y--axis, and the spatial order
  of the Chebyshev expansion of the spatial model along the x--axis.
  The {\it left} column shows these results for prefiltering, while
  the {\it right} column shows this for the post--filtered analysis.
  The best results listed in Table~\ref{tab-bestfp}, with the
  associated configurations from Table~\ref{tab-bestconfig},
  correspond to the minimum in each of the subpanels.
}
\label{fp_heatmap}
\end{figure}

@@

>>>>

\subsection{Algorithm and Background}

@@

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We use variations on the
configuration where we use one Gaussian, with the above width, and
three Gaussians, the central of which has this width, and the others a
factor of $\beta = 2.0$ smaller and larger.  We modify each of the
Gaussians by a set of Laguerre polynomials of a given order.  The
smallest Gaussian is modified to the highest order, with the others
modified by order//2.  Figure~\ref{basis} provides an example basis
set with this configuration.  The total number of shapes in the basis
are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription for basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 3.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.
}.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

A template image was generated by {\bf DETAIL}.  The template has a
seeing of 0.88''.  This means that the matching of the template image
with visit 6866601 will yield a template {\it deconvolution} in the
postfilter configuration.  The matching of the template with 6866601
in the prefilter configuration, and with 6866602 in the postfilter
configuration, will yield a small convolution, since both images will
have a FWHM of 0.8--0.9''.  All other Psf--matching operations will
yield a significant convolution of the template image.

@@

@@

\subsection{Background Subtraction}

\subsection{Number of False Positives With Detection Threshold}

To estimate the numbers of false positives we would expect in a
$4000\times4072$ pixel science image with a given FWHM in pixels, we
do the following.  The theoretical expectations for the number of
false positives due to random Gaussian fluctuations is outlined in
\cite{Kaiser-PointSources}.

\subsection{Variance Misestimation}

\subsection{Measurements on Pre--filtered Data}

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results: Production 8} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 5                      & 29  \\
         & False     & 10                     & 81  \\
v6866602 & True      & 5                      & 30  \\
         & False     & 9                      & 35  \\
v6866603 & True      & 2                      & 9   \\
         & False     & 1                      & 23  \\
\end{tabular}
\caption{Total number of {\tt DiaSources} detected at 5--sigma from
  all 9 sensors in raft 2,2, for the configurations that yielded the
  lowest number of false positives.  Matches are determined using a
  2'' match radius with the input reference catalog.  Objects with the
  {\tt EDGE} flag are ignored. \label{tab-bestfp8}}
\end{table}

\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results: Production 10} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 0                      & 70   \\
         & False     & 0                      & 143  \\
v6866602 & True      & 0                      & 42   \\
         & False     & 0                      & 53   \\
v6866603 & True      & 0                      & 23   \\
         & False     & 0                      & 36   \\
\end{tabular}
\caption{Total number of {\tt DiaSources} detected at 5--sigma from
  all 9 sensors in raft 2,2, for the configurations that yielded the
  lowest number of false positives.  Matches are determined using a
  2'' match radius with the input reference catalog.  Objects with the
  {\tt EDGE} flag are ignored. \label{tab-bestfp10}}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Configs: Production 8} \\
\hline
Visit    & Prefilter & nGauss & degGauss & spatialOrder \\
\hline
v6866601 & True      & 3      & 6        & 5 \\
         & False     & 3      & 2        & 4 \\
v6866602 & True      & 3      & 6        & 5 \\
         & False     & 3      & 5        & 5 \\
v6866603 & True      & 3      & 6        & 4 \\
         & False     & 3      & 6        & 4 \\
\end{tabular}
\caption{Configurations that led to the best results listed in
  Table~\ref{tab-bestfp8}.  All post-production runs include
  perturbations about these configurations. \label{tab-bestconfig8}}
\end{table}

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Configs: Production 10} \\
\hline
Visit    & Prefilter & nGauss & degGauss & spatialOrder \\
\hline
v6866601 & True      & 3      & 6        & 6 \\
         & False     & 3      & 1        & 5 \\
v6866602 & True      & 3      & 6        & 6 \\
         & False     & 3      & 6        & 5 \\
v6866603 & True      & 3      & 5        & 5 \\
         & False     & 3      & 4        & 6 \\
\end{tabular}
\caption{Configurations that led to the best results listed in
  Table~\ref{tab-bestfp10}.  All post-production runs include
  perturbations about these configurations. \label{tab-bestconfig10}}
\end{table}

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/heatmap10.eps} \\
\caption{``Heat--map'' detailing the total numbers of 5--sigma false
  positives as a function of two main degrees of freedom: the maximum
  order of basis Laguerre polynomials along the y--axis, and the
  spatial order of the Chebyshev expansion of the spatial model along
  the x--axis.  The {\it left} column shows these results for
  prefiltering, while the {\it right} column shows this for the
  post--filtered analysis.  The best results listed in
  Table~\ref{tab-bestfp}, with the associated configurations from
  Table~\ref{tab-bestconfig}, correspond to the minimum in each of the
  subpanels.  }
\label{fp_heatmap}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/cellsize.eps} \\
\caption{False positives with cell size
}
\label{cellsize}
\end{figure}

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/edge.eps} \\
\caption{Distribution of false positives as a function of the minimum
  distance from the edge of the sensor.  The {\it green} line shows
  the results from a random distribution of points.  The {\it blue}
  lines show the measured distributions, normalized as a probability
  density ($\sum y~\Delta x = 1$).  Error bars come from the square
  root of the number of points in each bin. }
\label{edgedist}
\end{figure}
